---
title: "Establish trimmings origins"
output: html_document
date: "2024-07-24"
---

## Summary

Combine origins information from Cargill, Skretting, BioMar, and other literature, aggregated in scripts in 02_compile_fm_origins folder.

    
When there are countries and species which are in multiple reports (e.g., Atlantic herring exists for Denmark for both Cargill and BioMar, but they report different trimmings proportions), we take the average proportion.


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(glue)
library(sf)
library(devtools)
library(janitor)
library(countrycode)
library(taxize)
library(qs)

source(here("R/directories.R"))

artis_genus_df <- read.csv(file.path(artis_dir, "artis_full_data/data/attribute tables/sciname.csv")) ## read in attribute table with common names


```



```{r}
biomar_int <- read.csv(here("int/material_origins_biomar_raw.csv")) %>%
  dplyr::select(-fao_fishing_area) %>% # dont need this
  mutate(
         common_name = ifelse(common_name == "pacific sardine", "south american pilchard", common_name)) %>% #  ## just missing "pacific sardine", which is sardinops sagax, common name is reported as south american pilchard in ARTIS.. we'll use what is listed in ARTIS for the name
  mutate(country = ifelse(country == "Other nei", "unknown", country),
         iso3c = ifelse(country == "unknown", "unknown", iso3c)) %>%
  mutate(sciname = tolower(sciname)) %>%
      mutate(common_name = ifelse(common_name %in% c("herrings, sardines nei"), "herrings", common_name)) %>% # fix this one
  dplyr::select(-sciname) %>%
  left_join(artis_genus_df, by = "common_name") %>% # keep artis scinames
  dplyr::select(1:13) %>%
  mutate(trash_fish = "no")

 
cargill_int <- read.csv(here("int/material_origins_cargill.csv")) %>%
  mutate(capture = ifelse(sciname == "Salmo salar", "no", "yes")) %>%  ## add capture column
  filter(!is.na(sciname)) %>%
  filter(!(common_name %in% c("Miscellaneous species", "Mixed fish"))) %>%
  mutate(common_name = tolower(common_name),
         sciname = tolower(sciname)) %>%
    mutate(common_name = case_when(
    sciname == "ammodytes sp." ~ "sandeels(=sandlances) nei",
    common_name %in% c("peruvian anchoveta") ~ "anchoveta(=peruvian anchovy)",
    common_name == "blue whiting" ~ "blue whiting(=poutassou)",
    common_name == "european pilchard" ~ "european pilchard(=sardine)", 
    common_name == "chub mackerel" ~ "pacific chub mackerel", 
    common_name == "alaska pollock" ~ "alaska pollock(=walleye poll.)",
    common_name == "sardine" ~ "sardinellas nei",
    common_name == "humboldt squid" ~ "jumbo flying squid",
    common_name == "thread herring" ~ "pacific thread herring",
    common_name == "pangasius" ~ "pangas catfishes nei",
    TRUE ~ common_name
  )) %>%
      left_join(artis_genus_df, by = "common_name") %>%
    dplyr::select(-sciname.x) %>%
  rename(sciname = sciname.y) %>%
  dplyr::select(1:13) %>%
    mutate(trash_fish = "no")




  
  
skretting_int <- read.csv(here("int/material_origins_skretting_global.csv")) %>%
    filter(!is.na(country)) %>% # ok skretting is just missing trimmings_tonnes and wholefish_tonnes, for now we'll leave as NA 
    mutate(trimmings_tonnes = NA, wholefish_tonnes = NA) %>%
  mutate(common_name = ifelse(sciname == "ammodytes tobianus", "small sandeel", common_name)) %>%
  mutate(common_name = ifelse(sciname == "ammodytes spp", "sandeels(=sandlances) nei", common_name)) %>%
  mutate(common_name = case_when(
    sciname == "ammodytes tobianus" ~ "small sandeel",
    sciname == "ammodytes spp" ~ "sandeels(=sandlances) nei",
    common_name == "sardine" ~ "sardinellas nei",
    sciname == "thunnini spp" ~ "tunas nei",
    common_name == "hake" ~ "hakes nei",
    common_name == "menhaden" ~ "gulf menhaden",
    common_name %in% c("chilean anchoveta", "peruvian anchoveta") ~ "anchoveta(=peruvian anchovy)",
    common_name == "cod" ~ "atlantic cod",
    common_name == "blue whiting" ~ "blue whiting(=poutassou)",
    common_name == "coho salmon" ~ "coho(=silver) salmon", 
    sciname == "pollachius pollachius" ~ "pollack",
    sciname == "pollachius virens" ~ "saithe(=pollock)",
    sciname == "prionotus carolinus" ~ "northern searobin",
    common_name == "european sardine" ~ "european pilchard(=sardine)", 
    common_name == "chub mackerel" ~ "pacific chub mackerel", 
    common_name == "orange-spotted spinefoot" ~ "goldlined spinefoot",
    common_name == "sprat" ~ "european sprat", 
    common_name == "jackmackerel" ~ "pacific jack mackerel",
    TRUE ~ common_name
  )) %>%
  left_join(artis_genus_df, by = "common_name") %>%
  dplyr::select(-sciname.x) %>%
  rename(sciname = sciname.y) %>%
  dplyr::select(1:13) %>%
    mutate(trash_fish = "no")

cao_int <- read.csv(here("int/material_origins_cao_chn.csv")) %>%
  dplyr::select(1:11, common_name) %>%
      mutate(trimmings_tonnes = NA, wholefish_tonnes = NA)

chiu_int <- read.csv(here("int/material_origins_chiu_chn.csv")) %>%
  mutate(trimmings_tonnes = NA, wholefish_tonnes = NA) %>%
  mutate(country = "China")

greenpeace_int <- read.csv(here("int/material_origins_greenpeace_chn.csv")) %>%
  dplyr::select(1:11, common_name) %>%
      mutate(trimmings_tonnes = NA, wholefish_tonnes = NA) 

peron_int <- read.csv(here("int/material_origins_peron_global.csv")) %>%
        mutate(trimmings_tonnes = NA, wholefish_tonnes = NA,
               country = countrycode(sourcevar = iso3c, origin = "iso3c", destination = "country.name")) 

  
all_int_dfs <- rbind(biomar_int, cargill_int, skretting_int, cao_int, chiu_int, greenpeace_int, peron_int) %>%
  mutate(sciname = tolower(sciname),
         common_name = tolower(common_name)) %>%
  mutate(capture = ifelse(iso3c == "MEX" & common_name == "yellowfin tuna", "no", capture)) # make sure mexico is listed as aquaculture. Oman also lists wholefish reduction, however, oman doesnt have any tuna aquaculture

write.csv(all_int_dfs, here("int/material_origins_ALL.csv"), row.names = FALSE)

```

Now lets summarise so that we don't have any duplicate countries and species. 

 - just take the regular mean
 
 
```{r}

int_dfs_summary <- all_int_dfs %>%
  group_by(sciname, common_name, iso3c, capture, trash_fish) %>%
  summarise(
    # trim_prop_w_mean = weighted.mean(trim_prop, trimmings_tonnes, na.rm = TRUE),
            trim_prop_mean = mean(trim_prop, na.rm = TRUE)) %>%
  ungroup() %>%
 # mutate(trim_prop_w_mean = ifelse(trim_prop_w_mean == "NaN", 0, trim_prop_w_mean)) %>% # fix those with 0 trimmings tonnes
#  mutate(trim_prop_w_mean = ifelse(is.na(trim_prop_w_mean), trim_prop_mean, trim_prop_w_mean)) %>%
  dplyr::select(sciname, common_name, iso3c, capture, trash_fish, trim_prop = trim_prop_mean) %>%
  mutate(trimmings = case_when(
    trim_prop == 1 ~ "yes",
    trim_prop > 0 & trim_prop < 1 ~ "yes",
    trim_prop == 0 ~ "no"
  ),
  wholefish = case_when(
    trim_prop == 1 ~ "no",
        trim_prop > 0 & trim_prop < 1 ~ "yes",
    trim_prop == 0 ~ "yes"
  )) 


write.csv(int_dfs_summary, here("outputs/all_fmfo_material_origins_reports.csv"), row.names = FALSE)

```
 

