---
title: "Establish trimmings origins"
output: html_document
date: "2024-07-24"
---

## Summary


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(glue)
library(sf)
library(devtools)
library(janitor)
library(countrycode)
library(taxize)
library(qs)

source(here("R/directories.R"))

full_artis <- qs::qread(file.path(artis_dir, "full_artis_fm_fillet.qs"))

fm_artis <- full_artis %>% 
  filter(hs6 == 230120)

fillet_artis <- full_artis %>%
  filter(hs6 != 230120)


artis_taxa <- read.csv(file.path(artis_dir, "artis_full_data/data/attribute tables/sciname.csv")) ## read in attribute table with common names
```

Testing: Join ARTIS dta to biomar and skretting information. Provide summary stats for how much info is missing. 


```{r}
fm_origins_df_fin <- read.csv(here("int/fishmeal_origins_biomar_skretting.csv")) %>%
  dplyr::select(sciname, fao_fishing_area, country, iso3c, wholefish, trimmings, trim_prop, capture, data_source_feed, data_source_fao_area_country, feed_report_spp_name)

artis_trimmings <- fm_artis %>%
  left_join(artis_taxa) %>%
#  filter(year == 2020) %>%
  left_join(fm_origins_df_fin, by = c("sciname", "source_country_iso3c" = "iso3c")) %>% ## some duplicates
  mutate(wholefish = ifelse(family == "salmonidae", "no", wholefish),
         trimmings = ifelse(family == "salmonidae", "yes", trimmings),
         trim_prop = ifelse(family == "salmonidae", 1, trim_prop)) %>%  # make all salmon trimmings no matter what 
  mutate(trimmings = ifelse(method == "aquaculture", "yes", trimmings), # make any aquaculture method trimmings no matter what? 
         wholefish = ifelse(method == "aquaculture", "no", wholefish),
         trim_prop = ifelse(method == "aquaculture", 1, trim_prop)) %>%
  mutate(capture = ifelse(method == "capture", "yes", capture)) %>%
  mutate(capture = ifelse(method == "aquaculture", "no", capture)) %>%
  mutate(trimmings = ifelse(str_detect(family, "scombridae"), "yes", trimmings),
         wholefish = ifelse(str_detect(family, "scombridae"), "no", wholefish),
         trim_prop = ifelse(str_detect(family, "scombridae"), 1, trim_prop)) %>% # # make any tuna catch trimmings 
  left_join(artis_taxa) %>%  # get taxanomic info
  distinct() %>%
mutate(trim_liveweight = trim_prop*live_weight_t,
    trim_product = trim_prop*product_weight_t, 
    wholefish_liveweight = (1-trim_prop)*live_weight_t,
    wholefish_product = (1-trim_prop)*product_weight_t) ## allocate with what info we have? 


test <- artis_trimmings %>% filter(is.na(common_name)) ## 0 ok this is good! 

## I think there are some duplicates, so i will need to see what is happening there. 

duplicates <- artis_trimmings %>%
  dplyr::select(-live_weight_t, -product_weight_t, -trim_liveweight, -trim_product) %>%
  group_by_all() %>%
  summarize(count = n()) %>%
  filter(count > 1) ## no duplicates!!!??
  

### lets do some exploration on what is missing! 

## check for 2020 

## Species and countries that we have NO info on if it is trimmings or wholefish reduction
test <- artis_trimmings %>%
  filter(year == 2020) %>% 
  filter(is.na(wholefish) & is.na(trimmings))

sum(test$product_weight_t) # 748903.1
sum(test$live_weight_t) # 2228142

total_product_weight <- artis_trimmings %>% 
  filter(year == 2020) %>%
  pull(product_weight_t) %>%
  sum() # 3697986

total_live_weight <- artis_trimmings %>% 
  filter(year == 2020) %>%
  pull(live_weight_t) %>%
  sum() # 11002272


748903.1/total_product_weight # 0.2025165 product
2228142/total_live_weight # 0.2025165 liveweight

## ok about a fifth of FM production in 2020, we have no info for! 



## trimmings only species
test <- artis_trimmings %>%
  filter(year == 2020) %>% 
  filter(wholefish == "no" & trimmings == "yes") 

sum(test$trim_product) # 941243.7
sum(test$trim_liveweight) # 2800394


941243.7/total_product_weight # 0.2545287 product
2800394/total_live_weight # 0.2545287 liveweight
## ok, ~25% we know for sure are used exclusively for trimmings... this includes crustaceans though, so I'm not actually confident in that 

## lets take out crustaceans from that estimate to see

# https://www.feedipedia.org/node/202#:~:text=Shrimp%20meal%20or%20shrimp%20waste,made%20of%20fresh%20water%20shrimps.
# "Shrimp meal or shrimp waste meal is the undecomposed ground dried waste of shrimp. There are several types of shrimp meal on the market depending on the kind of raw material used. It can contain whole shrimps and/or shrimp parts such as the heads or shells. Shrimp meal may also be made of fresh water shrimps. Shrimp meal is a by-product of the shrimp industry. Approximately 70% of total shrimp landings become waste, so there is a tremendous tonnage of shrimp waste produced. A portion of the crude protein that is contained in shrimp meal is in the form of chitin, which is not readily digestible (Göhl, 1982)."

## so it seems like shrimp could be "wholefish" or "trimmings". Need to do some research on this...

# https://www.researchgate.net/publication/375923791_Potential_industrial_and_nutritional_applications_of_shrimp_by-products_a_review#:~:text=The%20main%20by%2Dproducts%20include,%2C%20and%20anti%2Doxidant%20properties.
## This suggests that shrimp byproducts from processing make up 40-60% of the whole shrimp.. so seems likely that most fishmeal from shrimp would be from trimmings?

# https://www.globalseafood.org/advocate/shrimp-processing-byproducts-find-many-uses/ ; this suggests the same thing and specifically says the byproducts are used to make feeds

test <- artis_trimmings %>%
  filter(year == 2020) %>% 
  filter(wholefish == "no" & trimmings == "yes") %>%
  filter(phylum != "arthropoda")

sum(test$trim_product) # 739300.6
sum(test$trim_liveweight) # 2199572


739300.6/total_product_weight # 0.1999198 product
2199572/total_live_weight # 0.1999198 live weight





## species that can be wholefish or trimmings

test <- artis_trimmings %>%
  filter(year == 2020) %>% 
  filter(wholefish == "yes" & trimmings == "yes")

sum(test$trim_product) # 591726.4
sum(test$trim_liveweight) # 1760508



591726.4/total_product_weight # 0.1600132 product trimmings 
1760508/total_live_weight # 0.1600132 liveweight trimmings


sum(test$wholefish_product) # 1232488
sum(test$wholefish_liveweight) # 3666906

1232488/total_product_weight # 0.3332863 product trimmings 
3666906/total_live_weight # 0.3332863 liveweight trimmings


## so now we are estimating ~41% of FM comes from trimmings, not including any of the missing species we have. IFFO reports 32% for 2020
## without crustaceans included in that it is ~36%


## species and areas we know for sure are only wholefish

test <- artis_trimmings %>%
  filter(year == 2020) %>% 
  filter(wholefish == "yes" & trimmings == "no")

sum(test$product_weight_t) # 183624.7
sum(test$live_weight_t) # 546321.5



183624.7/total_product_weight # 0.04965684 product
546321.5/total_live_weight # 0.04965683 liveweight
## we know for sure wholefish is ~5% + 34% = 39%

## so if we believe IFFO, then most of the remaining 20% should be mostly wholefish rendered? These numbers seem high for trimmings! 

```


Right now we're matching based on iso3c and sciname. We could first gapfill by taxa first and then use price data. However, this would assume that trimmings inclusion is the same regardless of origin. 

Let's try that, with the average proportion inclusion. And if it is a yes once, then we assume it is a yes everywhere.

```{r}

missing_artis_info <- artis_trimmings %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings) & is.na(wholefish))

## gapfill the both trimmings and wholefish species. For this, we will assume that no matter where the species comes from, the inclusion rate is the same (an average... probably change this to weighted average or something)
# so if a species is listed as trimmings and wholefish in one place, we will gapfill if that species is traded as fishmeal elsewhere when we don't have info 

wholefish_trimmings_spp <- artis_trimmings %>%
    filter(year == 2020, 
           trimmings == "yes",
         wholefish == "yes") %>%
 # distinct(sciname, trim_prop, trimmings, wholefish,) %>%
  group_by(sciname, trimmings, wholefish) %>%
  summarise(trim_prop = weighted.mean(trim_prop, live_weight_t, na.rm = TRUE)) %>%
  ungroup()

gapfill_missing_1 <- missing_artis_info %>%
  dplyr::select(-trim_prop, -trimmings, -wholefish) %>%
  left_join(wholefish_trimmings_spp) %>%
  filter(!is.na(trimmings))
# ok that gapfilling only fixes 1448 observations

## lets do the same thing but with the wholefish only species 


wholefish_spp <- artis_trimmings  %>%
    filter(year == 2020, 
           trimmings == "no",
         wholefish == "yes") %>%
  # distinct(sciname, trim_prop, trimmings, wholefish) %>%
  group_by(sciname, trimmings, wholefish) %>%
  summarise(trim_prop = weighted.mean(trim_prop, live_weight_t, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(!(sciname %in% c(gapfill_missing_1$sciname)))

gapfill_missing_2 <- missing_artis_info %>%
  dplyr::select(-trim_prop, -trimmings, -wholefish) %>%
  left_join(wholefish_spp) %>%
  filter(!is.na(trimmings))

# ok that gapfilling only fixes 881 observations

## lets do the same thing but with the trimmings only species

trimmings_spp <- artis_trimmings  %>%
  filter(year == 2020) %>% 
    filter(trimmings == "yes",
         wholefish == "no") %>%
 # distinct(sciname, trim_prop, trimmings, wholefish) %>%
  group_by(sciname, trimmings, wholefish) %>%
  summarise(trim_prop = weighted.mean(trim_prop, live_weight_t, na.rm = TRUE)) %>%
  ungroup() %>%
    filter(!(sciname %in% c(gapfill_missing_1$sciname))) %>%
    filter(!(sciname %in% c(gapfill_missing_2$sciname)))

gapfill_missing_3 <- missing_artis_info %>%
  dplyr::select(-trim_prop, -trimmings, -wholefish) %>%
  left_join(trimmings_spp) %>%
  filter(!is.na(trimmings))

 # ok that gapfilling fixes 28846 observations

## so if we did the species level gapfilling first, as above, we would have ~75k to still gapfill with prices

gapfilling_spp_level <- rbind(gapfill_missing_1, gapfill_missing_2, gapfill_missing_3) # ok now we have ~29k more rows with info


missing_artis_info_new <- missing_artis_info %>% 
  filter(!(sciname %in% c(unique(gapfilling_spp_level$sciname)))) # filter those rows out of the original missing info

nrow(missing_artis_info_new) # 78979
nrow(gapfilling_spp_level) # 14355
14355 + 78979 # 93334 
nrow(missing_artis_info) # 93334 perfect!

full_artis_gapfill_spp <- rbind(missing_artis_info_new, gapfilling_spp_level) %>%
  rbind(., artis_trimmings %>%   filter(year == 2020, !is.na(trimmings) & !is.na(wholefish))) %>%
  mutate(trim_liveweight = trim_prop*live_weight_t,
    trim_product = trim_prop*product_weight_t, 
    wholefish_liveweight = (1-trim_prop)*live_weight_t,
    wholefish_product = (1-trim_prop)*product_weight_t) ## allocate with what info we have? 

test <- artis_trimmings %>% filter(year == 2020)
 
nrow(full_artis_gapfill_spp) == nrow(test) # TRUE - perfect! 


## now lets check the stats again

### lets do some exploration on what is missing! 

## check for 2020 

## Species and countries that we have NO info on if it is trimmings or wholefish reduction
test <- full_artis_gapfill_spp %>%
  filter(year == 2020) %>% 
  filter(is.na(wholefish) & is.na(trimmings))

lw <- sum(test$product_weight_t)
pw <- sum(test$live_weight_t) 

total_product_weight <- full_artis_gapfill_spp %>% 
  filter(year == 2020) %>%
  pull(product_weight_t) %>%
  sum() 

total_live_weight <- full_artis_gapfill_spp %>% 
  filter(year == 2020) %>%
  pull(live_weight_t) %>%
  sum() 


lw/total_product_weight # 0.1783523 product
pw/total_live_weight # 0.1783523 liveweight

## ok now down to 18% of FM production in 2020, we have no info for! 



## trimmings only species
test <- full_artis_gapfill_spp %>%
  filter(year == 2020) %>% 
  filter(wholefish == "no" & trimmings == "yes") 

pw = sum(test$trim_product) # 
lw = sum(test$trim_liveweight) # 


pw/total_product_weight # 0.2739991 product
lw/total_live_weight # 0.2739991 liveweight
## ok, ~26% we know for sure are used exclusively for trimmings... this includes crustaceans though, so I'm not actually confident in that 

## lets take out crustaceans from that estimate to see

test <- full_artis_gapfill_spp %>%
  filter(year == 2020) %>% 
  filter(wholefish == "no" & trimmings == "yes") %>%
  filter(phylum != "arthropoda")

pw = sum(test$trim_product) # 
lw = sum(test$trim_liveweight) # 


pw/total_product_weight # 0.2146011 product
lw/total_live_weight # 0.2146011 live weight





## species that can be wholefish or trimmings

test <- full_artis_gapfill_spp %>%
  filter(year == 2020) %>% 
  filter(wholefish == "yes" & trimmings == "yes")

pw = sum(test$trim_product) # 
lw = sum(test$trim_liveweight) # 



pw/total_product_weight # 0.1611577 product trimmings 
lw/total_live_weight # 0.1611577 liveweight trimmings


sum(test$wholefish_product) # 
sum(test$wholefish_liveweight) # 

1234089/total_product_weight # 0.3337192 product wholefish 
3671668/total_live_weight # 0.3337192 liveweight wholefish


## so now we are estimating ~43% of FM comes from trimmings, not including any of the missing species we have. IFFO reports 32% for 2020
## without crustaceans included in that it is ~37%


## species and areas we know for sure are only wholefish

test <- full_artis_gapfill_spp %>%
  filter(year == 2020) %>% 
  filter(wholefish == "yes" & trimmings == "no")

sum(test$product_weight_t) # 195149.4
sum(test$live_weight_t) # 580609.8



195149.4/total_product_weight # 0.05277181 product
580609.8/total_live_weight # 0.05277181 liveweight
## we know for sure wholefish is ~5% + 34% = 39%

## so if we believe IFFO, then most of the remaining 18% should be mostly wholefish rendered

```

Let's identify trimmed species by getting species which are filleted

```{r}

fillet_2020 <- fillet_artis %>% 
  filter(year == 2020) %>%
  distinct(source_country_iso3c, exporter_iso3c, importer_iso3c, sciname) %>%
  mutate(filleted = "yes")


missing_artis_fillet <- missing_artis_info_new %>%
  left_join(fillet_2020)



test <- missing_artis_fillet %>% 
  group_by(filleted) %>%
  summarise(n(),
            lw_sum = sum(live_weight_t))

## if we allocated all filleted species, then would add 6% more to trimmings.. so that would be ~49% from trimmings
678748.1/total_live_weight # 0.06169163

1283532.1/total_live_weight # 0.1166606
```


 Gapfilling: 
 - look into prices of fish to determine trimmings or wholefish rendered spp
  - e.g., a high value fish like tuna is not going to be rendered whole fish to fish meal (we already account for this above, but the example stands)
  - e.g., Lobster is high value, if there is FM from lobster, it is likely all trimmings

```{r}
## use price data to inform trimmings spp; for example, nobody will be catching snapper for wholefish rendering into FM.. if the fish is above a certain $$ ====> trimmings.. there is price data on rdsi specifically for FMFO.. Establish this cutoff price based on yearly prices. Can also incorporate fishmeal prices.. e.g. they wont render spp that exceed value of FM 
 # - join ts of FM to price data


prices <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/outputFMFO2021-03-12.csv")
taxa <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/taxa.csv")

entity <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/fishing_entity.csv") %>%
  dplyr::select(-X) %>%
  mutate(iso3c = countrycode(sourcevar = name, origin = "country.name", destination = "iso3c"))

prices_taxa <- prices %>% 
  left_join(., taxa, by = "TaxonKey") %>%
    clean_names() %>%
  left_join(., entity, by = "fishing_entity_id") %>%
  mutate(sciname = tolower(taxon_name),
         common_name = tolower(common_name)) %>%
  dplyr::select(source_country_iso3c = iso3c, sciname, common_name, iso3c, year, price2010usd_mean, price2010usd_ci) %>%
  filter(!is.na(source_country_iso3c)) %>% # filter out territories for now
  filter(year == 2016) %>% # filter for 2016 for testing
  dplyr::select(-year, -common_name)


missing_artis_info <- artis_trimmings %>%
  filter(is.na(wholefish) & is.na(trimmings)) %>%
  filter(year == 2020)


missing_taxa_prices <- missing_artis_info %>%
  # dplyr::select(sciname) %>%
  # distinct(sciname) %>%
  left_join(prices_taxa, by = c("sciname", "source_country_iso3c")) %>%
  distinct()
## ~half don't have price info.. so will need to use some sort of mean, or if available a price from a prior year 


## ok lets look at species which we know are wholefish ONLY and see what their prices are like

wholefish_only <- artis_trimmings %>%
  filter(wholefish == "yes",
         trimmings == "no", 
         year == 2020) %>%
  left_join(prices_taxa, by = c("sciname", "source_country_iso3c")) %>%
  distinct() %>%
  filter(common_name != "antarctic krill")

## missing prices for ~25%

## max price is $324 

## ok lets look at species which we know are both wholefish and trimmings and see what their prices are like

wholefish_trimmings_only <- artis_trimmings %>%
  filter(wholefish == "yes",
         trimmings == "yes", 
         year == 2020) %>%
  left_join(prices_taxa, by = c("sciname", "source_country_iso3c")) %>%
  distinct() %>%
  dplyr::select(source_country_iso3c, exporter_iso3c, importer_iso3c, sciname, method, trim_prop, common_name, price2010usd_mean)

summary(wholefish_trimmings_only)
## missing prices for ~20%

## pretty variable prices across trimmings inclusion rates
## max price is ~1000USD 


```
