---
title: "Explore trimmings origins"
output: html_document
date: "2024-08-01"
---

## Summary


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(glue)
library(sf)
library(devtools)
library(janitor)
library(countrycode)
library(qs)

source(here("R/directories.R"))

full_artis <- qs::qread(file.path(artis_dir, "full_artis_fm_fillet.qs"))

fm_artis <- full_artis %>% 
  filter(hs6 == 230120)

fillet_artis <- full_artis %>%
  filter(hs6 != 230120)


artis_taxa <- read.csv(file.path(artis_dir, "artis_full_data/data/attribute tables/sciname.csv")) ## read in attribute table with common names
countries <- read.csv(here("data/countries.csv")) ## artis countries with info


```

Join ARTIS data to biomar and skretting information. Provide summary stats for how much info is missing. 


```{r}
fm_origins_df_fin <- read.csv(here("outputs/all_fmfo_material_origins_reports.csv"))


ggplot(fm_origins_df_fin) +
  geom_histogram(aes(trim_prop)) +
  theme_bw() +
  labs(x = "Proportion of fishmeal derived from trimmings", 
       y = "Distinct species and source country count")

artis_trimmings_all <- fm_artis %>%
  left_join(artis_taxa) %>%
  left_join(fm_origins_df_fin, by = c("sciname", "source_country_iso3c" = "iso3c", "common_name")) 

artis_trimmings_info <- artis_trimmings_all %>% 
  filter(!is.na(trimmings)) %>%
  mutate(method_gf = NA,
         trimmings_gf = trimmings,
         wholefish_gf = wholefish,
         trim_prop_gf = trim_prop) %>%
  mutate(trimmings_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", "yes", trimmings_gf),
         wholefish_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", "no", wholefish_gf),
        trim_prop_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", 1, trim_prop_gf)) # fix the mexico yellowfin tuna instance

# test <- artis_trimmings_info %>% filter(is.na(wholefish)) # ok perfect, both empty
# test <- artis_trimmings_info %>% filter(is.na(trim_prop))

artis_trimmings_missing <- artis_trimmings_all %>% 
  filter(is.na(trimmings))

# nrow(artis_trimmings_missing) + nrow(artis_trimmings_missing) == nrow(artis_trimmings_all) # TRUE

# test <- artis_trimmings_missing %>% filter(!is.na(wholefish))
# test <- artis_trimmings_missing %>% filter(!is.na(trim_prop)) # perfect empty

artis_trimmings_gf <- artis_trimmings_missing %>% 
  mutate(trimmings_gf = 
           case_when(
             method == "aquaculture" ~ "yes",
             family == "salmonidae" ~ "yes",
             sciname == "salmoninae" ~ "yes",
             str_detect(family, "scombridae") ~ "yes",
             TRUE ~ trimmings
           ),
         wholefish_gf = 
           case_when(
             method == "aquaculture" ~ "no",
             family == "salmonidae" ~ "no",
             sciname == "salmoninae" ~ "no",
             str_detect(family, "scombridae") ~ "no",
             TRUE ~ wholefish
           ),
        trim_prop_gf = 
           case_when(
             method == "aquaculture" ~ 1, # assume aquaculture is all trimmings, unless feed reports suggest otherwise
             family == "salmonidae" ~ 1, # assume salmon is all trimmings
             sciname == "salmoninae" ~ 1, # assume salmon is all trimmings
             str_detect(family, "scombridae") ~ 1, # assume tuna is all trimmings
             TRUE ~ trim_prop
           ),
        method_gf = 
           case_when(
             method == "aquaculture" ~ "aquaculture", # assume aquaculture is all trimmings
             family == "salmonidae" ~ "salmon", # assume salmon is all trimmings
             sciname == "salmoninae" ~ "salmon", # assume salmon is all trimmings
             str_detect(family, "scombridae") ~ "tuna", # assume tuna is all trimmings
             TRUE ~ NA
           )
         ) %>%
  mutate(capture = 
           case_when(
             method == "capture" ~ "yes",
             method == "aquaculture" ~ "no",
             TRUE ~ capture
           )) %>%
  distinct()



## rbind back together after prelimnary gapfilling

artis_trimmings_all_gf <- artis_trimmings_info %>%
  rbind(., artis_trimmings_gf) %>%
  distinct() %>%
  dplyr::select(-wholefish, -trimmings, -trim_prop) %>%
mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

# test <- artis_trimmings_all_gf %>% 
#   filter(!is.na(trimmings_gf) & is.na(wholefish_gf))
# 
# test <- artis_trimmings_all_gf %>% 
#   filter(is.na(trimmings_gf) & !is.na(wholefish_gf))
# 
# test <- artis_trimmings_all_gf %>% 
#   filter(family == "scombridae")
# 
# 
# test <- artis_trimmings_all_gf %>% 
#   filter(method == "aquaculture") %>%
#   filter(genus == "thunnus")

## ok looks like all gapfilling went good

test <- artis_trimmings_all_gf %>% filter(is.na(common_name)) ## 0 ok this is good! 

## lets save this to RDSI

write.csv(artis_trimmings_all_gf, file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf.csv"), row.names = TRUE)

```

Data checking and exploration

Let's look at just finfish species and see how much is missing. We believe that IFFO might only report finfish destined for human consumption, which would explain why our trimmings numbers are larger.


```{r}

artis_trimmings_all_gf <- read.csv(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf.csv"))

colnames(artis_trimmings_all_gf)


```

Make some more pretty plots

```{r}

artis_all_2020 <- artis_trimmings_all_gf %>%
  filter(year == 2020) %>%
  left_join(countries, by = c("source_country_iso3c" = "iso3c")) %>%
  mutate(owid_region = ifelse(is.na(owid_region), "unknown", owid_region))

total_product_weight <- sum(artis_all_2020$product_weight_t, na.rm = TRUE) # 3494636


## Let's make a bar plot and have x axis be the type of fish meal, y axis the proportion it represents, and fill by continent

no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "Not sure") %>%
  group_by(fishmeal_type, owid_region) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "Trimmings", "Whole fish")) %>%
  group_by(fishmeal_type, owid_region) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) 

## oceania: #9F6144
## Africa: #A660A1
## Asia: #3F8D87
## South America: #8F464E
## North America: #E47669
## Europe: #5E74A1
## unknown: #808080
  
ggplot(bar_plot_df, aes(x = reorder(fishmeal_type, -prop), y = prop, fill = owid_region)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#A660A1", "#3F8D87", "#5E74A1", "#E47669", "#9F6144", "#8F464E", "#808080")) +
  theme_classic() +
  labs(fill = "Region", y = "Proportion of global trade", x = "Fishmeal type") + 
    scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) 




## Lets look at imports and exports

no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) %>%
  group_by(source_country_iso3c) %>%
  mutate(total_product_w_iso3c = sum(total, na.rm = TRUE)) %>%
  arrange(-total_product_w_iso3c) %>%
  ungroup()

test <- bar_plot_df %>%
  distinct(source_country_iso3c, total_product_w_iso3c) %>%
  arrange(-total_product_w_iso3c) 

test2 <- test[1:30,]

bar_plot_fin <- bar_plot_df %>%
  filter(source_country_iso3c %in% c(unique(test2$source_country_iso3c)))
  
  
ggplot(bar_plot_fin, aes(x = reorder(source_country_iso3c, -prop), y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(fill = "Fishmeal type", y = "Proportion of global trade", x = "Country", title = "Top 30 fishmeal producers") 
  
#+
 # theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))



no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, exporter_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, exporter_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) %>%
  group_by(exporter_iso3c) %>%
  mutate(total_product_w_iso3c = sum(total, na.rm = TRUE)) %>%
  arrange(-total_product_w_iso3c) %>%
  ungroup()

test <- bar_plot_df %>%
  distinct(exporter_iso3c, total_product_w_iso3c) %>%
  arrange(-total_product_w_iso3c) 

test2 <- test[1:30,]

bar_plot_fin <- bar_plot_df %>%
  filter(exporter_iso3c %in% c(unique(test2$exporter_iso3c)))
  
  
ggplot(bar_plot_fin, aes(x = reorder(exporter_iso3c, -prop), y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(fill = "Fishmeal type", y = "Proportion of global trade", x = "Country", title = "Top 30 fishmeal exporters") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```


Look at finfish only 

```{r}
artis_finfish <- artis_trimmings_all_gf %>%
  filter(year == 2020) %>% 
  filter(phylum %in% c("chordata", NA)) %>% # ok only finfish (or NA)
  filter(!(common_name %in% c("sea urchins", "saltwater mussels", "sea lamprey", "lampreys nei")))

total_product_weight <- artis_finfish %>% 
  pull(product_weight_t) %>%
  sum() # 2996610

no_info_df <- artis_finfish %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- artis_finfish %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

ggplot(circle_plot_df, aes(x = fishmeal_type, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") +
  # coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = prop-0.015, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(x = "Fishmeal type", y = "Proportion of global trade") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position="none"
  )
```


Cao et al 2015
 - Around 40% of China's domestically produced fish meal is from processing waste 
 - European Union forbids use of farmed fish by-products in finfish feeds but allows them to be used in crustacean diets or vice versa. 


Let's look at just China and see if trimmings represents 40% of its FM

```{r}

artis_chn <- artis_all_2020 %>%
  filter(source_country_iso3c == "CHN")


no_info_df <- artis_chn %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()


bar_plot_df <- artis_chn %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
     rbind(., no_info_df) %>%
  mutate(total_product_t = sum(total, na.rm = TRUE)) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%"))

## ok right now, we have ~64% of China's FM coming from trimmings. This is really high, so its probably safe to assume that the rest is from wholefish, especially considering their large trash fish market there. 

## lets gapfill based on fillets and see where the proportions end up 

fillet_chn <- fillet_artis %>% 
  filter(source_country_iso3c == "CHN") %>%
  pull(sciname) %>%
  unique() ## 127 spp

setdiff(unique(artis_chn$sciname), fillet_chn) # 61 
setdiff(fillet_chn, unique(artis_chn$sciname)) # 31



artis_fillet_chn <- artis_chn %>% 
  mutate(trim_prop_gf = ifelse(is.na(trim_prop_gf) & sciname %in% c(fillet_chn), 1, trim_prop_gf)) %>%
 # mutate(trim_prop_gf = ifelse(is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing


no_info_df <- artis_fillet_chn %>%
  filter(is.na(trim_prop_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()


bar_plot_df <- artis_fillet_chn %>% 
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
     rbind(., no_info_df) %>%
  mutate(total_product_t = sum(total, na.rm = TRUE)) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%"))

## ok so doing the fillet thing seems pretty unreasonable... estimating ~90% from trimmings this way! 

## maybe i need a list of trash fish species? 

```


```{r}

artis_trimmings_all_gf <- read.csv(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf.csv"))

### lets do some exploration on what is missing! 

## do some data checks for 2020 

artis_2020 <- artis_trimmings_all_gf %>% 
  filter(year == 2020)

## Species and countries that we have NO info on if it is trimmings or wholefish reduction
test1 <- artis_2020 %>%
  filter(is.na(wholefish_gf) & is.na(trimmings_gf))


sum(test1$product_weight_t) # 662046.4
sum(test1$live_weight_t) # 1969725

total_product_weight <- artis_2020 %>% 
  pull(product_weight_t) %>%
  sum() # 3494636

total_live_weight <- artis_2020 %>% 
  pull(live_weight_t) %>%
  sum() # 10397265


662046.4/total_product_weight # 0.1894464 product
1969725/total_live_weight # 0.1894464 liveweight

## ok about 19% of FM production in 2020, we have no info for! 


## make cirlce chart showing this 


no_info_df <- artis_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- artis_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```


Right now we're matching based on iso3c AND sciname. We could gapfill by taxa first and then use price data. However, this would assume that trimmings inclusion is the same regardless of origin. 

Let's try that gapfilling on sciname, regardless of origin, with the average proportion inclusion. And if it is a yes once, then we assume it is a yes everywhere.

For example, we will get an average trim_prop value for species, regardless of origin, and fill in the species. Should we do a weighted average on product weight? Yes. 

```{r}

sciname_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(sciname, common_name) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

spp_level_gapfill <- artis_trimmings_all_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(sciname_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_gf <- artis_trimmings_all_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., spp_level_gapfill)

nrow(full_artis_spp_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with species averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

```


Right now we're matching based on iso3c AND sciname. In the last chunk we gapfilled by species.

INSTEAD of that, let's try that gapfilling on source_iso3c, regardless of species, with the average proportion inclusion.

For example, we will get an average trim_prop value for iso3c, regardless of species, and fill in the values.

```{r}

iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_level_gapfill <- artis_trimmings_all_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_iso3c_gf <- artis_trimmings_all_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., iso3c_level_gapfill)

nrow(full_artis_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with source country averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

```

Ok, now lets gapfill the species level gapfill dataset with iso3c averages. So the gapfill flow is like this: 

 - species level first
 - iso3c level second
 
 
```{r}

iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_spp_level_gapfill <- full_artis_spp_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_iso3c_gf <- full_artis_spp_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., iso3c_spp_level_gapfill)

nrow(full_artis_spp_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect


## make cirlce chart showing this 


no_info_df <- full_artis_spp_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c( "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled by species and then source country averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Ok, so the above great and all, but I don't think it necessarily reflects reality. Using price data in conjunction with fillets is probably a better way to gapfill... then if we still need to gapfill after that, we can use species and/or country level averages

Let's pull species which report fillets. We could do a number of things:

 - For our missing info: any species that is filleted, assume it is ALL trimmings (except salmon/tuna/aquaculture). Any species which is NOT filleted is considered ALL whole fish (probably not right, as shrimp/prawns are probably "trimmed").


```{r}

fillet_data <- fillet_artis %>% 
  distinct(year, source_country_iso3c, exporter_iso3c, importer_iso3c, sciname) %>%
  mutate(filleted = "yes")


simple_fillet_gf <- artis_trimmings_all_gf %>%
  left_join(fillet_data) %>%
  mutate(
  trimmings_gf = case_when(
    filleted == "yes" & is.na(trimmings_gf) ~ "yes",
    is.na(filleted) & is.na(trimmings_gf) ~ "no",
    TRUE ~ trimmings_gf
  ), 
  wholefish_gf = case_when(
    filleted == "yes" & is.na(wholefish_gf) ~ "no",
    is.na(filleted) & is.na(wholefish_gf) ~ "yes",
    TRUE ~ wholefish_gf
  )
  ) %>%
  mutate(trim_prop_gf = case_when(
    wholefish_gf == "yes" & trimmings_gf == "no" & is.na(trim_prop_gf) ~ 0,
    wholefish_gf == "no" & trimmings_gf == "yes" & is.na(trim_prop_gf) ~ 1,
    TRUE ~ trim_prop_gf
  )) %>%
    mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) 

test <- simple_fillet_gf %>%
  filter(is.na(trim_prop_gf)) # ok makes sense 


## make cirlce chart showing this 


no_info_df <- simple_fillet_gf %>%
  filter(year == 2020, source_country_iso3c == "CHN") %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- simple_fillet_gf %>%
  filter(year == 2020, source_country_iso3c == "CHN") %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c( "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled by filleted species") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

```


Any species that is filleted, assume these are all species which are "trimmings only" (for species that don't already have info). Join with price data with comprehensive trimmings info, and get an average price for trimmings species (any species in data with info and trimmings == yes). Any species below that price will be considered all whole fish fish meal, UNLESS it is a filleted species (since we're assuming those are ALL trimmings). 

```{r}
## use price data to inform trimmings spp; for example, nobody will be catching snapper for wholefish rendering into FM.. if the fish is above a certain $$ ====> trimmings.. there is price data on rdsi specifically for FMFO.. Establish this cutoff price based on yearly prices. Can also incorporate fishmeal prices.. e.g. they wont render spp that exceed value of FM 
 # - join ts of FM to price data

prices <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/outputFMFO2021-03-12.csv")
taxa <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/taxa.csv")

entity <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/fishing_entity.csv") %>%
  dplyr::select(-X) %>%
  mutate(iso3c = countrycode(sourcevar = name, origin = "country.name", destination = "iso3c"))

prices_taxa <- prices %>% 
  left_join(., taxa, by = "TaxonKey") %>%
    clean_names() %>%
  left_join(., entity, by = "fishing_entity_id") %>%
  mutate(sciname = tolower(taxon_name),
         common_name = tolower(common_name)) %>%
  dplyr::select(year, source_country_iso3c = iso3c, sciname, iso3c, year, price2010usd_mean, price2010usd_ci) %>%
  filter(!is.na(source_country_iso3c)) 
#%>% # filter out territories for now
 # left_join(artis_taxa)

## gapfill the fillets first 
fillet_gf <- artis_trimmings_all_gf %>%
  left_join(fillet_data) %>%
  mutate(
  trimmings_gf = case_when(
    filleted == "yes" & is.na(trimmings_gf) ~ "yes",
    TRUE ~ trimmings_gf
  ), 
  wholefish_gf = case_when(
    filleted == "yes" & is.na(wholefish_gf) ~ "no",
    TRUE ~ wholefish_gf
  )
  ) %>%
  mutate(trim_prop_gf = case_when(
    wholefish_gf == "no" & trimmings_gf == "yes" & is.na(trim_prop_gf) ~ 1,
    TRUE ~ trim_prop_gf
  ))


avg_trim_prices <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  left_join(prices_taxa) %>%
  group_by(sciname) %>%
  mutate(avg_usd_sciname = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>% ## missing ~15% of prices
  group_by(genus) %>%
    mutate(avg_usd_genus = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>% # missing ~6%
  group_by(family) %>%
  mutate(avg_usd_family = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
    group_by(order) %>%
  mutate(avg_usd_order = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
      group_by(class) %>%
  mutate(avg_usd_class = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(phylum) %>%
  mutate(avg_usd_phylum = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(kingdom) %>%
    mutate(avg_usd_kingdom = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(avg_usd_gf = ifelse(is.na(price2010usd_mean), avg_usd_sciname, price2010usd_mean)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_genus, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_family, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_order, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_class, avg_usd_gf)) %>%
    mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_phylum, avg_usd_gf)) %>%
    mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_kingdom, avg_usd_gf)) %>%
  filter(trim_prop_gf == 1) %>%
  group_by(year, source_country_iso3c) %>%
  summarise(avg_year_iso3c_prices = mean(avg_usd_gf, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(source_country_iso3c) %>%
  mutate(avg_iso3c_price = mean(avg_year_iso3c_prices, na.rm = TRUE)) %>%
  ungroup()


global_avg_price <- 339.8605


fillet_price_empty_gf <- fillet_gf %>%
  filter(is.na(filleted)) %>%
  filter(is.na(trimmings_gf)) %>%
  left_join(avg_trim_prices) %>%
  mutate(global_avg_price = 339.8605) %>%
  left_join(prices_taxa) %>%
  mutate(trim_prop_gf = ifelse(price2010usd_mean < avg_year_iso3c_prices, 0, trim_prop_gf)) %>%
    mutate(trim_prop_gf = ifelse(price2010usd_mean < avg_iso3c_price & is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
      mutate(trim_prop_gf = ifelse(price2010usd_mean < global_avg_price & is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
  dplyr::select(colnames(fillet_gf))

summary(fillet_price_empty_gf)


fillet_price_all_gf <- fillet_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., fillet_price_empty_gf) %>%
  distinct() %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) 

nrow(fillet_price_all_gf) == nrow(fillet_gf) # perfect
nrow(fillet_price_all_gf) == nrow(artis_trimmings_all_gf) # perfect

## make cirlce chart showing this 


no_info_df <- fillet_price_all_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- fillet_price_all_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with fillet and price data") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

```


Now add species average gapfilling to the above


```{r}
sciname_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(sciname, common_name) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

spp_level_gapfill <- fillet_price_all_gf %>%
  filter(is.na(trim_prop_gf)) %>%
  left_join(sciname_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_gf <- fillet_price_all_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., spp_level_gapfill)


## make cirlce chart showing this 


no_info_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Whole fish and by-products, gapfilled with fillet and price data, then species averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Now add country averages to the above

```{r}
iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_level_gapfill <- full_artis_spp_gf %>%
  filter(is.na(trim_prop_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_iso3c_gf <- full_artis_spp_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., iso3c_level_gapfill)

nrow(full_artis_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with fillet and price data, \nthen species and iso3c averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Let's take a look at the distribution of the "not sure" category in our data

 - by continent
 - by species taxa grouping


```{r}

artis_trimmings <- read.csv(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf.csv"))
colnames(artis_trimmings)

unknown_fm <- artis_trimmings %>% 
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  dplyr::select(-X)

unique(unknown_fm$source_country_iso3c)
length(unique(unknown_fm$sciname))

test <- unknown_fm %>% 
  group_by(source_country_iso3c, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE),
            product_weight_t = sum(product_weight_t, na.rm = TRUE)) %>%
  ungroup()


## test out plotting family

# Summarize the total product_weight_t for each family
family_summary <- unknown_fm %>%
  group_by(class) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0)  # Filter out families with no weight

# Group smaller families into "Other"
# threshold <- 0.01 * sum(family_summary$total_product_weight)  # 1% threshold
# family_summary <- family_summary %>%
#   mutate(family = ifelse(total_product_weight < threshold, "Other", family)) %>%
#   group_by(family) %>%
#   summarise(total_product_weight = sum(total_product_weight))

# Create the pie chart
ggplot(family_summary, aes(x = "", y = total_product_weight, fill = class)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Fish meal with missing information",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm %>% filter(is.na(class))

test <- unknown_fm %>% filter(class == "actinopterygii") %>%
  group_by(sciname) %>% 
  summarise(product_w = sum(product_weight_t)) %>%
  ungroup() %>%
  mutate(prop = product_w/363945.9)

## try isscaap category


# Summarize the total product_weight_t for each family
family_summary <- unknown_fm %>%
  group_by(isscaap) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0)  # Filter out families with no weight

# Create the pie chart
ggplot(family_summary, aes(x = "", y = total_product_weight, fill = isscaap)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Fish meal with missing information",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm %>% filter(str_detect(common_name, "prawn|shrimp"))

## now try continent 

countries <- read.csv(here("data/countries.csv"))

unknown_fm_countries <- unknown_fm %>%
  left_join(countries, by = c("source_country_iso3c" = "iso3c"))



# Summarize the total product_weight_t for each family
summary <- unknown_fm_countries %>%
  group_by(owid_region) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0) %>%  # Filter out families with no weight
  mutate(owid_region = ifelse(is.na(owid_region), "Unknown", owid_region))

# Create the pie chart
ggplot(summary, aes(x = "", y = total_product_weight, fill = owid_region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(fill = "Continent",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm_countries %>% filter(is.na(owid_region))
unique(test$source_country_iso3c) # all unknown

```

