---
title: "Explore trimmings origins"
output: html_document
date: "2024-08-01"
editor_options: 
  chunk_output_type: console
---

## Summary


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(glue)
library(sf)
library(devtools)
library(janitor)
library(countrycode)
library(qs)

source(here("R/directories.R"))

full_artis <- qs::qread(file.path(artis_dir, "full_artis_fm_fillet.qs"))

artis_version <- "old" # set version you want to run (old or new data from Jessica - this is temporary until we get more details on the new data from Jessica)


fillet_artis <- full_artis %>%
  filter(hs6 != 230120) %>%
  dplyr::select(-hs_version) %>% # don't need this
  rename(consumer_iso3c = importer_iso3c) # just to make things consistent

if(artis_version == "old"){
fm_artis <- full_artis %>%
  filter(hs6 == 230120) %>%
    rename(consumer_iso3c = importer_iso3c) %>%
  dplyr::select(-hs_version)

  
}else{

fm_artis <- qs::qread(file.path(artis_dir, "full_artis_fm_2024_09_19.qs")) %>%
  mutate(hs6 = 230120,
         dom_source = ifelse(dom_source == "", consumption_type, dom_source)) %>%
  dplyr::select(year, source_country_iso3c, exporter_iso3c, consumer_iso3c, dom_source, hs6, sciname, habitat, method, product_weight_t, live_weight_t = consumption_t_capped) %>%
  mutate(exporter_iso3c = ifelse(source_country_iso3c == consumer_iso3c, source_country_iso3c, exporter_iso3c))
}

setdiff(colnames(fillet_artis), colnames(fm_artis))
setdiff(colnames(fm_artis), colnames(fillet_artis)) # 0 good


artis_taxa <- read.csv(file.path(artis_dir, "artis_full_data/data/attribute tables/sciname.csv")) ## read in attribute table with common names
countries <- read.csv(here("data/countries.csv")) ## artis countries with info

asia_iso3c <- countries %>% 
  filter(owid_region == "Asia") %>%
  pull(iso3c)
```

Join ARTIS data to biomar and skretting information. Provide summary stats for how much info is missing. 


```{r}
fm_origins_df_fin <- read.csv(here("outputs/all_fmfo_material_origins_reports.csv"))


ggplot(fm_origins_df_fin) +
  geom_histogram(aes(trim_prop)) +
  theme_bw() +
  labs(x = "Proportion of fishmeal derived from trimmings", 
       y = "Distinct species and source country count")

artis_trimmings_all <- fm_artis %>%
  left_join(artis_taxa) %>%
  left_join(fm_origins_df_fin, by = c("sciname", "source_country_iso3c" = "iso3c", "common_name")) 

artis_trimmings_info <- artis_trimmings_all %>% 
  filter(!is.na(trimmings)) %>%
  mutate(method_gf = NA,
         trimmings_gf = trimmings,
         wholefish_gf = wholefish,
         trim_prop_gf = trim_prop) %>%
  mutate(trimmings_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", "yes", trimmings_gf),
         wholefish_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", "no", wholefish_gf),
        trim_prop_gf = ifelse(source_country_iso3c == "MEX" & common_name == "yellowfin tuna", 1, trim_prop_gf))  # fix the mexico yellowfin tuna instance

## For China/South East Asia I need to look into these papers; so if these prove true, then any NEI spceies in SE Asia will be "trash fish" and therefore whole fish reduction:

# https://www.science.org/doi/10.1126/sciadv.adn5650 say much trash fish fish meal is classified as “mixed fish” or “nei” in FAO
#         - sources: D. Leadbitter, “Driving change in South East Asian trawl fisheries, fishmeal supply, and aquafeed” [Report to IFFO, The Marine Ingredients Organisation and the Global Aquaculture Alliance (GAA), Fish Matter, 2019].
#         - U. R. Sumaila, W. W. L. Cheung, L. S. L. Teh, A. H. Y. Bang, T. Cashion, Z. Zeng, J. J. Alava, S. Le Clue, Y. Sadovy de Mitcheson, “Sink or swim: The future of fisheries in the East and South China Seas” (ADM Capital Foundation, 2021).
#         - Y. Sadovy de Mitcheson, D. Leadbitter, C. Law, “History, profiles and implications of feed fish and fishmeal supply from domestic trawlers in the East and South China Seas - Final Report to ADMCF” (2018).


# Majluf et al suggests trimmings are really only possible in large scale in  China, Norway, Chile, US, and Peru. This is not true. THA and VNM both have massive FM industries and produces a lot of trimmings MF. Does Morocco have large scale FM processing plants? Is there a list with large scale FM plant operations? Seems a reasonable assumption to make that if there is large scale industry present, then trimmings is more likely than not. 

# Leadbitter says that Pangasius is used for trimmings in VNM. Probably the same everywhere else. 

## do some testing around the China data
# test <- artis_trimmings_all %>%
#   filter(trash_fish == "yes", method == "aquaculture") # so there are no trash fish and aquaculture.. cool. all trash fish are capture. makes sense
# test <- artis_trimmings_all %>%
#   filter(method == "aquaculture", source_country_iso3c == "CHN") # this is all tilapia and all trimmings. Makes sense.

artis_trimmings_missing <- artis_trimmings_all %>% 
  filter(is.na(trimmings))

# nrow(artis_trimmings_missing) + nrow(artis_trimmings_missing) == nrow(artis_trimmings_all) # TRUE

## arthropoda and mollusc fish meal - what to do about this? It makes sense that their by-products would be used for FM, but it really makes results seem suspicious compared to IFFO reports, because IFFO only includes finfish.

artis_trimmings_gf <- artis_trimmings_missing %>% 
  mutate(trimmings_gf = 
           case_when(
             method != "aquaculture" & str_detect(common_name, "nei|ray-finned fishes") & phylum == "chordata" & source_country_iso3c %in% c(asia_iso3c) ~ "no", # start with china for this - will probably expand to all SE asian countries
                method != "aquaculture" & sciname == "animalia" ~ "no", # any unknown capture fisheries will be whole fish
             method == "aquaculture" & phylum == "chordata" ~ "yes", # if it is finfish aquaculture, then assume it is trimmings.
             family == "salmonidae" ~ "yes",
             sciname == "salmoninae" ~ "yes",
             str_detect(common_name, "tuna|Tuna|albacore|Albacore") ~ "yes", # any tuna species are trimmings
             TRUE ~ trimmings
           ),
         wholefish_gf = 
           case_when(
      method != "aquaculture" & str_detect(common_name, "nei|ray-finned fishes") & phylum == "chordata" & source_country_iso3c %in% c(asia_iso3c) ~ "yes",   
       method != "aquaculture" & sciname == "animalia" ~ "yes",
            method == "aquaculture" & phylum == "chordata" ~ "no",
             family == "salmonidae" ~ "no",
             sciname == "salmoninae" ~ "no",
             str_detect(common_name, "tuna|Tuna|albacore|Albacore") ~ "no", # any tuna species are trimmings
             TRUE ~ wholefish
           ),
        trim_prop_gf = 
           case_when(
                   method != "aquaculture" & str_detect(common_name, "nei|ray-finned fishes") & phylum == "chordata" & source_country_iso3c %in% c(asia_iso3c) ~ 0,   
                    method != "aquaculture" & sciname == "animalia" ~ 0,
             method == "aquaculture" & phylum == "chordata" ~ 1, # assume finfish aquaculture is all trimmings, unless feed reports suggest otherwise
             family == "salmonidae" ~ 1, # assume salmon is all trimmings
             sciname == "salmoninae" ~ 1, # assume salmon is all trimmings
             str_detect(common_name, "tuna|Tuna|albacore|Albacore") ~ 1, # assume tuna is all trimmings
             TRUE ~ trim_prop
           ),
        method_gf = 
           case_when(
                   method != "aquaculture" & str_detect(common_name, "nei|ray-finned fishes") & phylum == "chordata" & source_country_iso3c %in% c(asia_iso3c) ~ "trash fish",   
                    method != "aquaculture" & sciname == "animalia" ~ "unknown species, assume wholefish",
            method == "aquaculture" & phylum == "chordata" ~ "aquaculture", # assume finfish aquaculture is all trimmings
             family == "salmonidae" ~ "salmon", # assume salmon is all trimmings
             sciname == "salmoninae" ~ "salmon", # assume salmon is all trimmings
             str_detect(common_name, "tuna|Tuna|albacore|Albacore") ~ "tuna", # assume tuna is all trimmings
             TRUE ~ NA
           )
         ) %>%
  mutate(capture = 
           case_when(
             method == "capture" ~ "yes",
             method == "aquaculture" ~ "no",
             TRUE ~ capture
           )) %>%
  distinct()



test <- artis_trimmings_gf %>%
  filter(source_country_iso3c == "CHN", phylum == "chordata", method != "aquaculture", is.na(trim_prop_gf))

test <- artis_trimmings_gf %>%
  filter(sciname == "animalia")

## rbind back together after prelimnary gapfilling

artis_trimmings_all_gf <- artis_trimmings_info %>%
  rbind(., artis_trimmings_gf) %>%
  distinct() %>%
  dplyr::select(-wholefish, -trimmings, -trim_prop) %>%
mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

# test <- artis_trimmings_all_gf %>%
#   filter(!is.na(trimmings_gf) & is.na(wholefish_gf))
# 
# test <- artis_trimmings_all_gf %>% 
#   filter(is.na(trimmings_gf) & !is.na(wholefish_gf))
# 
# test <- artis_trimmings_all_gf %>% 
#   filter(family == "scombridae")
# 
# 
# test <- artis_trimmings_all_gf %>%
#   filter(method == "aquaculture") %>%
#   filter(genus == "thunnus")

## ok looks like all gapfilling went good


## lets save this to RDSI
if(artis_version == "old"){
qs::qsave(artis_trimmings_all_gf, file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf_traded.qs"))
}else{
  qs::qsave(artis_trimmings_all_gf, file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf_traded_consumption.qs"))

}

```

Data checking and exploration

Let's look at just finfish species and see how much is missing. We believe that IFFO might only report finfish destined for human consumption, which would explain why our trimmings numbers are larger.


```{r}

artis_trimmings_all_gf <- qs::qread(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf_traded.qs"))

colnames(artis_trimmings_all_gf)


```

Make some more pretty plots

```{r}

artis_all_2020 <- artis_trimmings_all_gf %>%
  filter(year == 2020) %>%
  left_join(countries, by = c("source_country_iso3c" = "iso3c")) %>%
  mutate(owid_region = ifelse(is.na(owid_region), "unknown", owid_region))

total_product_weight <- sum(artis_all_2020$product_weight_t, na.rm = TRUE) 
# 3494636
# 6158133 using new data with domestic consumption!? Not sure if that is right... FAO SOFIA says there is ~5 million tonnes

test <- artis_all_2020 %>%
  group_by(dom_source) %>%
  summarise(sum_t = sum(product_weight_t, na.rm = TRUE))

test <- artis_all_2020 %>%
  mutate(fm_consumed_source = ifelse(source_country_iso3c != consumer_iso3c, "domestic", "foreign")) %>%
  mutate(fm_consumed_source = ifelse(dom_source == "error", "error", fm_consumed_source)) %>%
  group_by(fm_consumed_source) %>%
  summarise(sum_t = sum(product_weight_t, na.rm = TRUE)) # interesting... claims about the same comes from domestic as foreign.. not sure that makes sense. 


test <- artis_all_2020 %>%
  filter(trim_product > 0) # yeah ok, the assumption that all aquaculture is trimmings is screwing things up big time. Because there is SO MUCH shrimps and prawns fish meal reported... look into this. 

test <- artis_all_2020 %>% 
  filter(is.na(trimmings_gf)) # let's look at whats missing. A lot of shrimp and bivalve. A lot of the finfish that are missing are larger groups like "nei", so I think if we assume the same rules as China for all of south east asia (VNM, THA, etc), this will help a lot for whole fish. 

## Let's make a bar plot and have x axis be the type of fish meal, y axis the proportion it represents, and fill by continent

no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "Not sure") %>%
  group_by(fishmeal_type, owid_region) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
 # mutate(fishmeal_type = ifelse(isscaap == "Shrimps, prawns", "wholefish_product", fishmeal_type)) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "Trimmings", "Whole fish")) %>%
  group_by(fishmeal_type, owid_region) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) 

## oceania: #9F6144
## Africa: #A660A1
## Asia: #3F8D87
## South America: #8F464E
## North America: #E47669
## Europe: #5E74A1
## unknown: #808080
  
ggplot(bar_plot_df, aes(x = reorder(fishmeal_type, -prop), y = prop, fill = owid_region)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#A660A1", "#3F8D87", "#5E74A1", "#E47669", "#9F6144", "#8F464E", "#808080")) +
  theme_classic() +
  labs(fill = "Region", y = "Proportion of global trade", x = "Fishmeal type") + 
    scale_x_discrete(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) # oh gosh... with the new data now ~50% from trimmings WITHOUT gapfilling...




## Lets look at imports and exports

no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) %>%
  group_by(source_country_iso3c) %>%
  mutate(total_product_w_iso3c = sum(total, na.rm = TRUE)) %>%
  arrange(-total_product_w_iso3c) %>%
  ungroup()

test <- bar_plot_df %>%
  distinct(source_country_iso3c, total_product_w_iso3c) %>%
  arrange(-total_product_w_iso3c) 

test2 <- test[1:30,]

bar_plot_fin <- bar_plot_df %>%
  filter(source_country_iso3c %in% c(unique(test2$source_country_iso3c)))
  
  
ggplot(bar_plot_fin, aes(x = reorder(source_country_iso3c, -prop), y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(fill = "Fishmeal type", y = "Proportion of global trade", x = "Country", title = "Top 30 fishmeal producers") +
theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) ## ok breaking down by country we see that its asia that is really contributing to trimmings... which tbh might not be THAT unreasonable...



no_info_df <- artis_all_2020 %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, exporter_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

bar_plot_df <- artis_all_2020 %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, exporter_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type))  %>%
  filter(prop > 0) %>%
  arrange(-prop) %>%
  group_by(exporter_iso3c) %>%
  mutate(total_product_w_iso3c = sum(total, na.rm = TRUE)) %>%
  arrange(-total_product_w_iso3c) %>%
  ungroup()

test <- bar_plot_df %>%
  distinct(exporter_iso3c, total_product_w_iso3c) %>%
  arrange(-total_product_w_iso3c) 

test2 <- test[1:30,]

bar_plot_fin <- bar_plot_df %>%
  filter(exporter_iso3c %in% c(unique(test2$exporter_iso3c)))
  
  
ggplot(bar_plot_fin, aes(x = reorder(exporter_iso3c, -prop), y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") + 
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(fill = "Fishmeal type", y = "Proportion of global trade", x = "Country", title = "Top 30 fishmeal exporters") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```


Look at finfish only 

```{r}
artis_finfish <- artis_trimmings_all_gf %>%
  filter(year == 2020) %>% 
  filter(phylum %in% c("chordata", NA)) %>% # ok only finfish (or NA)
  filter(!(common_name %in% c("sea urchins", "saltwater mussels", "sea lamprey", "lampreys nei")))

test <- artis_finfish %>%
  filter(is.na(trimmings_gf))

sum(test$product_weight_t) # 331945.7 missing total

## look at countries with trimmings:

test <- artis_finfish %>%
  filter(trimmings_gf == "yes") %>%
  group_by(source_country_iso3c) %>%
  summarise(total_trim_iso3c = sum(trim_product)) %>%
  ungroup() %>%
  mutate(total_trim_prop = total_trim_iso3c/sum(artis_finfish$trim_product, na.rm = TRUE)) %>%
  ungroup()
  
# Morocco is 1, unknown 2, Then PER, NOR, CHL; Majluf et al suggests trimmings are really only possible in large scale in  China, Norway, Chile, US, and Peru. Does Morocco have large scale FM processing plants? Is there a list with large scale FM plant operations? Seems a reasonable assumption to make that if there is large scale industry present, then trimmings is more likely than not. 
unique(test$source_country_iso3c) # lots of countries here 

total_product_weight <- artis_finfish %>% 
  pull(product_weight_t) %>%
  sum() # 2996927

no_info_df <- artis_finfish %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- artis_finfish %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

ggplot(circle_plot_df, aes(x = fishmeal_type, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", color = "white") +
  # coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = prop-0.015, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_bw() +
  labs(x = "Fishmeal type", y = "Proportion of global trade") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position="none"
  )

## seems a bit more reasonable/matching to IFFO looking only at finfish...
```


Cao et al 2015
 - Around 40% of China's domestically produced fish meal is from processing waste 
 - European Union forbids use of farmed fish by-products in finfish feeds but allows them to be used in crustacean diets or vice versa. 


Let's look at just China and see if trimmings represents 40% of its FM

```{r}

artis_chn <- artis_all_2020 %>%
  filter(source_country_iso3c == "CHN")


no_info_df <- artis_chn %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()


bar_plot_df <- artis_chn %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
     rbind(., no_info_df) %>%
  mutate(total_product_t = sum(total, na.rm = TRUE)) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%"))

## ok right now, we have ~70% of China's FM coming from trimmings. This is really high, so its probably safe to assume that the rest is from wholefish, especially considering their large trash fish market there. 

## lets gapfill based on fillets and see where the proportions end up 

fillet_chn <- fillet_artis %>% 
  filter(source_country_iso3c == "CHN") %>%
  pull(sciname) %>%
  unique() ## 127 spp

setdiff(unique(artis_chn$sciname), fillet_chn) # 61 
setdiff(fillet_chn, unique(artis_chn$sciname)) # 31



artis_fillet_chn <- artis_chn %>% 
  mutate(trim_prop_gf = ifelse(is.na(trim_prop_gf) & sciname %in% c(fillet_chn), 1, trim_prop_gf)) %>%
 # mutate(trim_prop_gf = ifelse(is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing


no_info_df <- artis_fillet_chn %>%
  filter(is.na(trim_prop_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()


bar_plot_df <- artis_fillet_chn %>% 
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type, source_country_iso3c) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
     rbind(., no_info_df) %>%
  mutate(total_product_t = sum(total, na.rm = TRUE)) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%"))

## ok so doing the fillet thing seems pretty unreasonable... estimating ~80% from trimmings this way! 

## maybe i need a list of trash fish species? 

```


```{r}

artis_trimmings_all_gf <- read.csv(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf_2024-10-17.csv"))

### lets do some exploration on what is missing! 

## do some data checks for 2020 

artis_2020 <- artis_trimmings_all_gf %>% 
  filter(year == 2020)

## Species and countries that we have NO info on if it is trimmings or wholefish reduction
test1 <- artis_2020 %>%
  filter(is.na(wholefish_gf) & is.na(trimmings_gf))


sum(test1$product_weight_t) # 662046.4
sum(test1$live_weight_t) # 1969725

total_product_weight <- artis_2020 %>% 
  pull(product_weight_t) %>%
  sum() # 3494636

total_live_weight <- artis_2020 %>% 
  pull(live_weight_t) %>%
  sum() # 10397265

```


Right now we're matching based on iso3c AND sciname. We could gapfill by taxa first and then use price data. However, this would assume that trimmings inclusion is the same regardless of origin. 

Let's try that gapfilling on sciname, regardless of origin, with the average proportion inclusion. And if it is a yes once, then we assume it is a yes everywhere.

For example, we will get an average trim_prop value for species, regardless of origin, and fill in the species. Should we do a weighted average on product weight? Yes. 

```{r}

sciname_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(sciname, common_name) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

spp_level_gapfill <- artis_trimmings_all_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(sciname_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_gf <- artis_trimmings_all_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., spp_level_gapfill)

nrow(full_artis_spp_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with species averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )
## just adds more to trimmings
```


Right now we're matching based on iso3c AND sciname. In the last chunk we gapfilled by species.

INSTEAD of that, let's try that gapfilling on source_iso3c, regardless of species, with the average proportion inclusion.

For example, we will get an average trim_prop value for iso3c, regardless of species, and fill in the values.

```{r}

iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_level_gapfill <- artis_trimmings_all_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_iso3c_gf <- artis_trimmings_all_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., iso3c_level_gapfill)

nrow(full_artis_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with source country averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )
# yikes
```

Ok, now lets gapfill the species level gapfill dataset with iso3c averages. So the gapfill flow is like this: 

 - species level first
 - iso3c level second
 
 
```{r}

iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_spp_level_gapfill <- full_artis_spp_gf %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_iso3c_gf <- full_artis_spp_gf %>%
  filter(!is.na(wholefish_product)) %>%
  rbind(., iso3c_spp_level_gapfill)

nrow(full_artis_spp_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect


## make cirlce chart showing this 


no_info_df <- full_artis_spp_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c( "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled by species and then source country averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Ok, so the above great and all, but I don't think it necessarily reflects reality. Using price data in conjunction with fillets is probably a better way to gapfill... then if we still need to gapfill after that, we can use species and/or country level averages

Let's pull species which report fillets. We could do a number of things:

 - For our missing info: any species that is filleted, assume it is ALL trimmings (except salmon/tuna/aquaculture). Any species which is NOT filleted is considered ALL whole fish (probably not right, as shrimp/prawns are probably "trimmed").


```{r}

fillet_data <- fillet_artis %>% 
  distinct(year, source_country_iso3c, exporter_iso3c, consumer_iso3c, sciname) %>%
  mutate(filleted = "yes")


simple_fillet_gf <- artis_trimmings_all_gf %>%
  left_join(fillet_data) %>%
  mutate(
  trimmings_gf = case_when(
    filleted == "yes" & is.na(trimmings_gf) ~ "yes",
    is.na(filleted) & is.na(trimmings_gf) ~ "no",
    TRUE ~ trimmings_gf
  ), 
  wholefish_gf = case_when(
    filleted == "yes" & is.na(wholefish_gf) ~ "no",
    is.na(filleted) & is.na(wholefish_gf) ~ "yes",
    TRUE ~ wholefish_gf
  )
  ) %>%
  mutate(trim_prop_gf = case_when(
    wholefish_gf == "yes" & trimmings_gf == "no" & is.na(trim_prop_gf) ~ 0,
    wholefish_gf == "no" & trimmings_gf == "yes" & is.na(trim_prop_gf) ~ 1,
    TRUE ~ trim_prop_gf
  )) %>%
    mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) 

test <- simple_fillet_gf %>%
  filter(is.na(trim_prop_gf)) # ok makes sense 


## make cirlce chart showing this 


no_info_df <- simple_fillet_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- simple_fillet_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c( "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled by filleted species") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )

# a bit more in line with reality (we think)

```


Any species that is filleted, assume these are all species which are "trimmings only" (for species that don't already have info). Join price data with comprehensive trimmings info, and get an average price for trimmings species (any species in data with info and trimmings == yes). Any species below that price will be considered all whole fish fish meal, UNLESS it is a filleted species (since we're assuming those are ALL trimmings). 

```{r}
## use price data to inform trimmings spp; for example, nobody will be catching snapper for wholefish rendering into FM.. if the fish is above a certain $$ ====> trimmings.. there is price data on rdsi specifically for FMFO.. Establish this cutoff price based on yearly prices. Can also incorporate fishmeal prices.. e.g. they wont render spp that exceed value of FM 
 # - join ts of FM to price data

prices <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/outputFMFO2021-03-12.csv")
taxa <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/taxa.csv")

entity <- read.csv("/mnt/rdsi/raw_data/sea-around-us/ex-vessel-prices/fishing_entity.csv") %>%
  dplyr::select(-X) %>%
  mutate(iso3c = countrycode(sourcevar = name, origin = "country.name", destination = "iso3c"))

prices_taxa <- prices %>% 
  left_join(., taxa, by = "TaxonKey") %>%
    clean_names() %>%
  left_join(., entity, by = "fishing_entity_id") %>%
  mutate(sciname = tolower(taxon_name),
         common_name = tolower(common_name)) %>%
  dplyr::select(year, source_country_iso3c = iso3c, sciname, iso3c, year, price2010usd_mean, price2010usd_ci) %>%
  filter(!is.na(source_country_iso3c)) 
#%>% # filter out territories for now
 # left_join(artis_taxa)

## gapfill the fillets first 
fillet_gf <- artis_trimmings_all_gf %>%
  left_join(fillet_data) %>%
  mutate(
  trimmings_gf = case_when(
    filleted == "yes" & is.na(trimmings_gf) ~ "yes",
    TRUE ~ trimmings_gf
  ), 
  wholefish_gf = case_when(
    filleted == "yes" & is.na(wholefish_gf) ~ "no",
    TRUE ~ wholefish_gf
  )
  ) %>%
  mutate(trim_prop_gf = case_when(
    wholefish_gf == "no" & trimmings_gf == "yes" & is.na(trim_prop_gf) ~ 1,
    TRUE ~ trim_prop_gf
  ))


avg_trim_prices <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  left_join(prices_taxa) %>%
  group_by(sciname) %>%
  mutate(avg_usd_sciname = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>% ## missing ~15% of prices
  group_by(genus) %>%
    mutate(avg_usd_genus = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>% # missing ~6%
  group_by(family) %>%
  mutate(avg_usd_family = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
    group_by(order) %>%
  mutate(avg_usd_order = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
      group_by(class) %>%
  mutate(avg_usd_class = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(phylum) %>%
  mutate(avg_usd_phylum = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(kingdom) %>%
    mutate(avg_usd_kingdom = mean(price2010usd_mean, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(avg_usd_gf = ifelse(is.na(price2010usd_mean), avg_usd_sciname, price2010usd_mean)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_genus, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_family, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_order, avg_usd_gf)) %>%
  mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_class, avg_usd_gf)) %>%
    mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_phylum, avg_usd_gf)) %>%
    mutate(avg_usd_gf = ifelse(is.na(avg_usd_gf), avg_usd_kingdom, avg_usd_gf)) %>%
  filter(trim_prop_gf == 1) %>%
  group_by(year, source_country_iso3c) %>%
  summarise(avg_year_iso3c_prices = mean(avg_usd_gf, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(source_country_iso3c) %>%
  mutate(avg_iso3c_price = mean(avg_year_iso3c_prices, na.rm = TRUE)) %>%
  ungroup()


global_avg_price <- mean(avg_trim_prices$avg_year_iso3c_prices)


fillet_price_empty_gf <- fillet_gf %>%
  filter(is.na(filleted)) %>%
  filter(is.na(trimmings_gf)) %>%
  left_join(avg_trim_prices) %>%
  mutate(global_avg_price = global_avg_price) %>%
  left_join(prices_taxa) %>%
  mutate(trim_prop_gf = ifelse(price2010usd_mean < avg_year_iso3c_prices, 0, trim_prop_gf)) %>%
    mutate(trim_prop_gf = ifelse(price2010usd_mean < avg_iso3c_price & is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
      mutate(trim_prop_gf = ifelse(price2010usd_mean < global_avg_price & is.na(trim_prop_gf), 0, trim_prop_gf)) %>%
  dplyr::select(colnames(fillet_gf))

summary(fillet_price_empty_gf)


fillet_price_all_gf <- fillet_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., fillet_price_empty_gf) %>%
  distinct() %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) 

nrow(fillet_price_all_gf) == nrow(fillet_gf) # perfect
nrow(fillet_price_all_gf) == nrow(artis_trimmings_all_gf) # perfect

## make cirlce chart showing this 


no_info_df <- fillet_price_all_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- fillet_price_all_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with fillet and price data") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )
## i guess there are still a lot of fish which are missing price info?
```


Now add species average gapfilling to the above


```{r}
sciname_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(sciname, common_name) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

spp_level_gapfill <- fillet_price_all_gf %>%
  filter(is.na(trim_prop_gf)) %>%
  left_join(sciname_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_spp_gf <- fillet_price_all_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., spp_level_gapfill)


## make cirlce chart showing this 


no_info_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_spp_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FF9999", "#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Whole fish and by-products, gapfilled with fillet and price data, then species averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Now add country averages to the above

```{r}
iso3c_averages <- artis_trimmings_all_gf %>%
  filter(!is.na(trimmings_gf)) %>%
  group_by(source_country_iso3c) %>%
  summarise(trim_prop_avg = weighted.mean(trim_prop_gf, product_weight_t)) %>%
  ungroup()

iso3c_level_gapfill <- full_artis_spp_gf %>%
  filter(is.na(trim_prop_gf)) %>%
  left_join(iso3c_averages) %>%
  mutate(trim_prop_gf = ifelse(!is.na(trim_prop_avg), trim_prop_avg, trim_prop_gf),
         method_gf = ifelse(!is.na(trim_prop_avg), "species", method_gf)) %>%
  mutate(trimmings_gf = case_when(
    trim_prop_gf == 1 ~ "yes",
    trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "no"
  ),
  wholefish_gf = case_when(
    trim_prop_gf == 1 ~ "no",
        trim_prop_gf > 0 & trim_prop_gf < 1 ~ "yes",
    trim_prop_gf == 0 ~ "yes"
  )) %>%
  dplyr::select(-trim_prop_avg) %>%
  mutate(trim_liveweight = trim_prop_gf*live_weight_t,
    trim_product = trim_prop_gf*product_weight_t, 
    wholefish_liveweight = (1-trim_prop_gf)*live_weight_t,
    wholefish_product = (1-trim_prop_gf)*product_weight_t) ## allocate with what info we have first for testing

## now join with un-gapfilled data

full_artis_iso3c_gf <- full_artis_spp_gf %>%
  filter(!is.na(trim_prop_gf)) %>%
  rbind(., iso3c_level_gapfill)

nrow(full_artis_iso3c_gf) == nrow(artis_trimmings_all_gf) # TRUE - perfect



## make cirlce chart showing this 


no_info_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf)) %>%
  mutate(fishmeal_type = "not sure") %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(product_weight_t)) %>%
  ungroup()

circle_plot_df <- full_artis_iso3c_gf %>%
  filter(year == 2020) %>% 
  filter(!is.na(trimmings_gf)) %>%
  pivot_longer(names_to = "fishmeal_type",
               cols = c("wholefish_product", "trim_product")) %>%
  mutate(fishmeal_type = ifelse(fishmeal_type == "trim_product", "trimmings", "wholefish")) %>%
  group_by(fishmeal_type) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rbind(., no_info_df) %>%
  mutate(total_product_t = total_product_weight) %>%
  mutate(prop = total/total_product_t) %>%
  mutate(perc = prop * 100) %>% # Convert to percentage
  mutate(label = paste0(round(perc, 1), "%")) %>%
  arrange(desc(fishmeal_type)) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)


# Plot the donut chart
ggplot(circle_plot_df, aes(x = 2, y = prop, fill = fishmeal_type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y", start = 0) +
  geom_text(aes(y = ypos, label = label), color = "black") +
 scale_fill_manual(values = c("#FFCC99", "#99CCFF")) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = "Fishmeal Type",
       title = "World's fishmeal in 2020",
       subtitle = "Gapfilled with fillet and price data, \nthen species and iso3c averages") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  )


```

Let's take a look at the distribution of the "not sure" category in our data

 - by continent
 - by species taxa grouping


```{r}

artis_trimmings <- read.csv(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_not_gf_2024-10-17.csv"))
colnames(artis_trimmings)

unknown_fm <- artis_trimmings %>% 
  filter(year == 2020) %>%
  filter(is.na(trimmings_gf) & is.na(wholefish_gf)) %>%
  dplyr::select(-X)

unique(unknown_fm$source_country_iso3c)
length(unique(unknown_fm$sciname))

test <- unknown_fm %>% 
  group_by(source_country_iso3c, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE),
            product_weight_t = sum(product_weight_t, na.rm = TRUE)) %>%
  ungroup()


## test out plotting family

# Summarize the total product_weight_t for each family
family_summary <- unknown_fm %>%
  group_by(class) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0)  # Filter out families with no weight

# Group smaller families into "Other"
# threshold <- 0.01 * sum(family_summary$total_product_weight)  # 1% threshold
# family_summary <- family_summary %>%
#   mutate(family = ifelse(total_product_weight < threshold, "Other", family)) %>%
#   group_by(family) %>%
#   summarise(total_product_weight = sum(total_product_weight))

# Create the pie chart
ggplot(family_summary, aes(x = "", y = total_product_weight, fill = class)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Fish meal with missing information",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm %>% filter(is.na(class))

test <- unknown_fm %>% filter(class == "actinopterygii") %>%
  group_by(sciname) %>% 
  summarise(product_w = sum(product_weight_t)) %>%
  ungroup() %>%
  mutate(prop = product_w/363945.9)

## try isscaap category


# Summarize the total product_weight_t for each family
family_summary <- unknown_fm %>%
  group_by(isscaap) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0)  # Filter out families with no weight

# Create the pie chart
ggplot(family_summary, aes(x = "", y = total_product_weight, fill = isscaap)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Fish meal with missing information",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm %>% filter(str_detect(common_name, "prawn|shrimp"))

## now try continent 

countries <- read.csv(here("data/countries.csv"))

unknown_fm_countries <- unknown_fm %>%
  left_join(countries, by = c("source_country_iso3c" = "iso3c"))



# Summarize the total product_weight_t for each family
summary <- unknown_fm_countries %>%
  group_by(owid_region) %>%
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE)) %>%
  filter(total_product_weight > 0) %>%  # Filter out families with no weight
  mutate(owid_region = ifelse(is.na(owid_region), "Unknown", owid_region))

# Create the pie chart
ggplot(summary, aes(x = "", y = total_product_weight, fill = owid_region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(fill = "Continent",
       x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "right")

test <- unknown_fm_countries %>% filter(is.na(owid_region))
unique(test$source_country_iso3c) # all unknown

### mostly unknowns from Asia

```

