---
title: "Save final data for analysis"
output: html_document
date: "2025-07-31"
editor_options: 
  chunk_output_type: console
---

## Summary

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(glue)
library(qs)
library(dplyr)
library(stringr)

source(here("R/directories.R"))

artis_taxa <- read.csv(file.path(artis_dir, "artis_full_data/data/attribute tables/sciname.csv")) %>% ## read in attribute table with common names
  mutate(phylum = case_when(
    common_name %in% c("mackerels, tunas, and bonitos", "salmons and trouts", "herrings", "cartilaginous fishes nei", "flounders", "african lungfishes", "toothfish", "carps, minnows, loaches, etc", "blue whitings") ~ "chordata",
    TRUE ~ phylum
  )) # fix phylum mistakes

fm_data <-   qs::qread(file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_traded_consumption_gf_final.qs")) %>%
  left_join(artis_taxa)

```

Adjust fishmeal data to include production and blood losses for trimmings weight

 - [Cao et al 2015](https://www.science.org/doi/10.1126/science.1260149): "We assume a median production yield for whole raw fish of 40% and a 2% flesh loss due to spoilage and other technical reasons." 
 - [Stevens et al. 2017](https://www.sciencedirect.com/science/article/pii/S0308597X17305328?via%3Dihub) referring to Atlantic Salmon: "For example, BPs in the form of trimmings represent a 2% yield of the whole fish"
 - [Sandstrom et al. 2022](https://www.nature.com/articles/s43016-022-00589-6): "The amounts of potential fish by-products were estimated by multiplying the amounts of fish for human consumption from capture fisheries and aquaculture by the ratios processed75,76 and multiplying the processed quantities by the average ratio of 41.5% of fish consisting of by-products (such as trimmings)79, subtracting 2% blood that is not used in reduction and finally assuming 2% losses at the primary fish processing stage"; 41.5% and 2% blood loss is based on Atlantic salmon, and the 2% processing is from Cao et al.
 
Based on this, we will assume that for finfish and non-bilvalve molluscs (i.e. squids and octopus), a 2% blood loss and 2% processing loss, everything else (namely shrimp/prawns) just a 2% processing loss (since they don't have blood). We are currently using an ~34% yield (i.e. 34% of the liveweight fish ends up rendered down to fishmeal) (this is the yield from ARTIS)


```{r}
fm_data_adjusted <- fm_data %>%
      mutate(phylum = case_when(is.na(phylum) & sciname %in% c("lampetra fluviatilis", "petromyzontidae", "petromyzon marinus") ~ "chordata",
                              is.na(phylum) & sciname == "animalia" ~ "animalia",
                            TRUE ~ phylum)) %>% 
    filter(phylum %in% c("chordata", "arthropoda") | phylum == "mollusca" & class == "cephalopoda") %>%
    mutate(trimmings = live_weight_t*trim_prop_gf,
         wholefish = live_weight_t*(1-trim_prop_gf)) %>%
  dplyr::select(-product_weight_t, -live_weight_t, -capture) %>%
  pivot_longer( 
    cols = c(trimmings, wholefish),
    names_to = "fishmeal_type",
    values_to = "live_weight_t"
  ) %>%
  mutate(product_weight_t =
           case_when(
             phylum == "arthropoda" & fishmeal_type == "trimmings" ~ (live_weight_t*0.98)/2.975207,
             phylum %in% c("mollusca", "chordata") & fishmeal_type == "trimmings" ~ (live_weight_t*0.96)/2.975207,
             TRUE ~ live_weight_t/2.975207
           )) 

qs::qsave(fm_data_adjusted, file.path(artis_dir, "artis_exchange/artis_fishmeal_trimmings_traded_consumption_adjusted_final.qs"))

test <- fm_data_adjusted %>%
  filter(year == 2020, 
         sciname == "engraulis ringens")

```

Save a file for top 20 contributing species per country for average years 2009-2013 (to match IFFO newton & jackson) and 2015-2019 (for latest years, excluding covid)

```{r}

  trade_flows_average <- fm_data_adjusted %>%
    filter(year %in% c(2009:2013)) %>%
    group_by(year, source_country_iso3c, exporter_iso3c, importer_iso3c = consumer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = sum(product_weight_t, na.rm = TRUE),
            live_weight_t = sum(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup() %>%
        group_by(source_country_iso3c, exporter_iso3c, importer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = mean(product_weight_t, na.rm = TRUE),
            live_weight_t = mean(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup()
  
  sum(trade_flows_average$product_weight_t) # 5442000 makes sense
  
top_species_by_country <- trade_flows_average %>%
  group_by(source_country_iso3c, sciname) %>% 
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE), .groups = "drop") %>%
  arrange(source_country_iso3c, desc(total_product_weight)) %>%
  group_by(source_country_iso3c) %>%
  slice_head(n = 20) %>%
  ungroup() %>%
  right_join(trade_flows_average, by = c("source_country_iso3c", "sciname")) %>%
    filter(!is.na(total_product_weight)) %>%
    left_join(artis_taxa) 
  
  sum(top_species_by_country$product_weight_t) # 4952253 ; should be less and it is

  ## save
  qs::qsave(top_species_by_country, here("outputs/top_species_fm_2009_2013.qs"))
  
  
  ## now do the same for 2015-2019
  
  trade_flows_average <- fm_data_adjusted %>%
    filter(year %in% c(2015:2019)) %>%
    group_by(year, source_country_iso3c, exporter_iso3c, importer_iso3c = consumer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = sum(product_weight_t, na.rm = TRUE),
            live_weight_t = sum(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup() %>%
        group_by(source_country_iso3c, exporter_iso3c, importer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = mean(product_weight_t, na.rm = TRUE),
            live_weight_t = mean(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup()
  
  sum(trade_flows_average$product_weight_t) # 5456924 makes sense
  
  
top_species_by_country <- trade_flows_average %>%
  group_by(source_country_iso3c, sciname) %>% 
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE), .groups = "drop") %>%
  arrange(source_country_iso3c, desc(total_product_weight)) %>%
  group_by(source_country_iso3c) %>%
  slice_head(n = 20) %>%
  ungroup() %>%
  right_join(trade_flows_average, by = c("source_country_iso3c", "sciname")) %>%
    filter(!is.na(total_product_weight)) %>%
    left_join(artis_taxa) 
  
  sum(top_species_by_country$product_weight_t) # 5000665 ; should be less and it is

  ## save
  qs::qsave(top_species_by_country, here("outputs/top_species_fm_2015_2019.qs"))
  
  
  
## Now lets save a 2009-2013 dataset WITH molluscs included and see what we get
  
trade_flows_average <-  fm_data %>%
      mutate(phylum = case_when(is.na(phylum) & sciname %in% c("lampetra fluviatilis", "petromyzontidae", "petromyzon marinus") ~ "chordata",
                              is.na(phylum) & sciname == "animalia" ~ "animalia",
                            TRUE ~ phylum)) %>% 
    filter(phylum %in% c("chordata", "arthropoda", "mollusca")) %>%
    mutate(trimmings = live_weight_t*trim_prop_gf,
         wholefish = live_weight_t*(1-trim_prop_gf)) %>%
  dplyr::select(-product_weight_t, -live_weight_t, -capture) %>%
  pivot_longer( 
    cols = c(trimmings, wholefish),
    names_to = "fishmeal_type",
    values_to = "live_weight_t"
  ) %>%
  mutate(product_weight_t =
           case_when(
             phylum == "arthropoda" & fishmeal_type == "trimmings" ~ (live_weight_t*0.98)/2.975207,
             phylum %in% c("mollusca", "chordata") & fishmeal_type == "trimmings" ~ (live_weight_t*0.96)/2.975207,
             TRUE ~ live_weight_t/2.975207
           )) %>%
       filter(year %in% c(2009:2013)) %>%
    group_by(year, source_country_iso3c, exporter_iso3c, importer_iso3c = consumer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = sum(product_weight_t, na.rm = TRUE),
            live_weight_t = sum(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup() %>%
        group_by(source_country_iso3c, exporter_iso3c, importer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = mean(product_weight_t, na.rm = TRUE),
            live_weight_t = mean(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup()
  
  sum(trade_flows_average$product_weight_t) # 7486643 makes sense
  
  
top_species_by_country <- trade_flows_average %>%
  group_by(source_country_iso3c, sciname) %>% 
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE), .groups = "drop") %>%
  arrange(source_country_iso3c, desc(total_product_weight)) %>%
  group_by(source_country_iso3c) %>%
  slice_head(n = 20) %>%
  ungroup() %>%
  right_join(trade_flows_average, by = c("source_country_iso3c", "sciname")) %>%
    filter(!is.na(total_product_weight)) %>%
    left_join(artis_taxa) 
  
  sum(top_species_by_country$product_weight_t) # 6811738 

  ## save
  qs::qsave(top_species_by_country, here("outputs/top_species_fm_2009_2013_inc_molluscs.qs"))

## now save every year of data 1996-2020 with only top 20 species
fm_data_top_spp <- fm_data_adjusted %>%
    group_by(year, source_country_iso3c, exporter_iso3c, importer_iso3c = consumer_iso3c, sciname, common_name, fishmeal_type, phylum) %>%
    summarise(
      product_weight_t = sum(product_weight_t, na.rm = TRUE),
            live_weight_t = sum(live_weight_t, na.rm = TRUE)

    ) %>%
    ungroup() %>%
  group_by(source_country_iso3c, sciname) %>% 
  summarise(total_product_weight = sum(product_weight_t, na.rm = TRUE), .groups = "drop") %>%
  arrange(source_country_iso3c, desc(total_product_weight)) %>%
  group_by(source_country_iso3c) %>%
  slice_head(n = 20) %>% # take the top 20 species for each source country (and year?)
  ungroup() %>%
  right_join(fm_data_adjusted, by = c("source_country_iso3c", "sciname")) %>%
    filter(!is.na(total_product_weight)) %>%
    left_join(artis_taxa) # this takes awhile to run

qs::qsave(fm_data_top_spp, here("outputs/top_species_fm_1996_2020.qs"))
  
```

